# Regression and model validation

This week I learning quite a bit about data wrangling and performing some basic linear regression

```{r}
date()
```
### Creating an enviroment with needed packages
```{r}
library(tidyverse)
library(GGally)
```
### Reading the data in and checking its structure
```{r}
students2014 <- read.csv("C:/Users/nicol/Downloads/PHD_302/IODS-project/learning2014.csv")
str(students2014)
```
### Creating a graphical overview of the data
```{r}
ggpairs(students2014, columns = 3:8)
```




#### Here I can see that most of the scores are normally distributed and with no real correlations that I can see


### Selecting variables to test a regression model

```{r}
first_model <- lm(formula= Points ~ attitude + deep  + Age  + surf + stra, data=students2014)
summary(first_model)
```


#### I honestly do not know enough about statistics to be able to say anything meaningful from this model.  I think that attitude seems to have the biggest impact on points while Age seems to have no effect, so I iwll remove it and see the what happens.


```{r}
second_model <- lm(formula= Points ~ attitude + deep  + surf + stra, data=students2014)
summary(second_model)
```


#### Here it seems that the impact of attitude increased when I took out Age, and the Multiple R squared went down.  My understanding of statistics is limited, but if the multiple R squared value went down (which ranges between 0 and 1, with 1 being a perfect correlation) the fact that it went down indicates the model fit better with Age included.


### Creating Diagnostic Plots

```{r}
plot(first_model, which = 1, caption = "Residuals vs Fitted")
plot(first_model, which = 2)
plot(first_model, which = 5)
```


#### *From the online textbook I see that linear regresssion models have 4 assumptions: Linear relationship between predictor variables and outcome, independence of variables, normal distribution of variables and equals variance of variables.

#### **Looking at the Residuals vs Fitted plot it looks like Points are fairly evenly distributed along their axis which indidcates that the models assumptions are valid

#### ***The Q-Q plot also seems to show a normal distriubution in residuals which is also a good check of the model

#### ****The leverage plot seems to show that not many residuals have a large impact on Points.  There are a few outliers that might merit removal.